getwd()
setwd('C:/Users/akshi/Downloads')
library(rmarkdown)
library(ggplot2)
library(dplyr)
library(ggvis)
library(Hmisc)
library(gridExtra)
library(tidyr)

Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size
Notes:
fb <- read.csv('pseudo_facebook.tsv', sep = '\t')
fb

***

### Scatterplots
Notes:

```{r Scatterplots}
qplot(x = fb$age, y = fb$friend_count, data = fb)
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}
summary(fb$age)
ggplot(aes(x = fb$age, y = fb$friend_count), data = fb) + geom_point() +
  xlim(13, 90)

```

***

### Overplotting
Notes:

```{r Overplotting}
ggplot(aes(x = fb$age, y = fb$friend_count), data = fb) + geom_point() +
  geom_point(alpha = 1/20) +
  xlim(13, 90)

ggplot(aes(x = fb$age, y = fb$friend_count), data = fb) + geom_point() +
  geom_jitter(alpha = 1/20) +
  xlim(13, 90)
```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}
plot <- ggplot(aes(x = fb$age, y = fb$friend_count), data = fb) +
  geom_point(alpha = 1/20) +
  xlim(13, 90)
plot
p1 <- plot + coord_trans(x = 'sqrt')
p1
p2 <- plot + coord_trans(y = 'sqrt')
p2

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes:

```{r Alpha and Jitter}
plot <- ggplot(aes(x = fb$age, y = fb$friendships_initiated), data = fb)

p1 <- plot + geom_point(alpha = 1/20 ) + xlim(13, 113)
p1
p2 <- plot + geom_jitter(alpha = 1/20) + xlim(13, 113)
p2

help("coord_cartesian")
p3 <- p1 + coord_cartesian(xlim = c(13, 113))
p3

```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
help(dplyr)
age <- fb$age
age_groups <- group_by(fb, age)
fb.fc_by_age <- summarise(age_groups)
fb.fc_by_age
fb.fc_by_age <- summarise(age_groups,
                          friend_count_mean = mean(friend_count),
                          friend_count_median = median(friend_count),
                          n = n())

fb.fc_by_age
fb.fc_by_age <- arrange(fb.fc_by_age, age)
fb.fc_by_age

head(fb.fc_by_age)

```

Create your plot!

```{r Conditional Means Plot}
fb.mean_fc <- mean(fb$friend_count)
fb.mean_fc
friend_count_mean = mean(friend_count)
plot <- ggplot(aes(x = age, y = friend_count_mean), data = fb.fc_by_age)
p1 <- plot + geom_line()
p1
p2 <- plot + geom_point()
p2

```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
friend_count_mean = mean(friend_count)
friend_count_mean
plot <- ggplot(aes(x = age, y = friend_count_mean), data = fb) +
  xlim(13, 113) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'yellow') +
  coord_trans(y = 'sqrt') +
  geom_line(stat = 'summary', fun.y = mean)
plot

plot <- plot + geom_line()

```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:

```{r Strong Correlations}

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

