Lesson 6
========================================================
getwd()
setwd('C:/Users/akshi/Downloads')
#install.packages('rmarkdown')
library(rmarkdown)
#install.packages('ggplot2')
library(ggplot2)
#install.packages('dplyr')
library(dplyr)
#install.packages('ggvis')
library(ggvis)
#install.packages('Hmisc')
library(Hmisc)
#install.packages('gridExtra')
library(gridExtra)
#install.packages('tidyr')
library(tidyr)


### Welcome
Notes:

***

### Scatterplot Review

```{r Scatterplot Review}
data(diamonds)
names(diamonds)
colnames(diamonds)

p1 <- ggplot(aes(diamonds$carat, y = diamonds$price),
             data = diamonds) +
  geom_point(fill = I('#E56810'), color = I('blue'), shape = 21)

p1

#using quantile function to eliminate the top 1 percent of the data
p1.1 <- p1 +
  coord_cartesian(xlim = c(0, quantile(diamonds$carat, 0.99)),
                  ylim = c(0, quantile(diamonds$price, 0.99)))
p1.1

#stat_smooth function to add a linear regression model to represent the data
p1.2 <- p1.1 +
  stat_smooth(method = 'lm')
p1.2

#showing three plots in a row
grid.arrange(p1, p1.1, p1.2)
```

***

### Price and Carat Relationship
Response:

***

### Frances Gerety
Notes:

#### A diamonds is
# forever

***

### The Rise of Diamonds
Notes:

***

### ggpairs Function
Notes:

```{r ggpairs Function}
# install these if necessary
#Ggally for ggpairs
#install.packages('GGally')
#install.packages('scales')
#memisc to summarise the linear regression
#install.packages('memisc')
#install.packages('lattice')
#install.packages('MASS')
#car to recode variables
#install.packages('car')
#to reshape and wrangle the data
#install.packages('reshape')
# plyr to create interesting summaries of the data
#install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp, 
        lower = list(continuous = wrap("points", shape = I('.'))), 
        upper = list(combo = wrap("box", outlier.shape = I('.'))))

```

What are some things you notice in the ggpairs output?
Response:

***

### The Demand of Diamonds
Notes:

```{r The Demand of Diamonds}
p1 <- ggplot(data = diamonds,
             aes(x = diamonds$price)) +
  geom_histogram(fill = I('#F00001'), binwidth = 100) +
  ggtitle('Price')
p1

p2 <- ggplot(data = diamonds,
             aes(x = diamonds$price)) +
  geom_histogram(fill = I('#F00001'), binwidth = 0.01) + 
  scale_x_log10() +
  ggtitle('Price(Log 10)')
p2

grid.arrange(p1, p2, nrow = 1)
```

***

### Connecting Demand and Price Distributions
Notes:

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
p3 <- ggplot(data = diamonds,
             aes(x = diamonds$carat, y = diamonds$price)) +
  geom_point(color = I('green')) +
  scale_y_log10() #+
#  stat_smooth(method = 'lm')
p3

```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', 
                                      transform = function(x) x^(1/3), 
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(x = diamonds$carat, y = diamonds$price),
       data = diamonds) + 
  geom_point() +
  scale_x_continuous(trans = cuberoot_trans(), 
                     limits = c(0.2, 4),
                     breaks = c(0.2, 0.5, 1, 2, 3, 4)) + 
  scale_y_continuous(trans = log10_trans(), 
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')

```

***

### Overplotting Revisited

```{r Sort and Head Tables}
carat.table <- table(diamonds$carat)
sort.ct.tbl <- sort(carat.table, decreasing = T)
head(sort.ct.tbl)

head(sort(table(diamonds$price), decreasing = T))

```
#a descending sorted head view of the tabled variables in the diamonds data set
# revelead that there are a huge number of data for the same dependent variable
# As this would cause overplotting the same data
# we use alpha to jitter and make the plot transparent
# to a certain extent

```{r Overplotting Revisited}

ggplot(aes(diamonds$carat, diamonds$price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 0.75, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), 
                     limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), 
                     limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')


```

***

### Other Qualitative Factors
Notes:

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
#install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = diamonds$carat, y = diamonds$price, color = diamonds$clarity), 
       data = diamonds) + 
  geom_point(alpha = 0.5, 
             size = 1, 
             position = 'jitter') +
  scale_color_brewer(type = 'div', 
                     guide = guide_legend(title = 'Clarity', 
                                          reverse = T, 
                                          override.aes = list(alpha = 1, size = 2)
                                          )
                     ) +  
  scale_x_continuous(trans = cuberoot_trans(), 
                     limits = c(0.2, 3), 
                     breaks = c(0.2, 0.5, 1, 2, 3)
                     ) + 
  scale_y_continuous(trans = log10_trans(), 
                     limits = c(350, 15000), 
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response:

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = diamonds$carat, y = diamonds$price, color = diamonds$cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Cut and Price
Response:

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}

ggplot(aes(x = diamonds$carat, y = diamonds$price, color = diamonds$color), 
       data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```

***

### Color and Price
Response:

***

### Linear Models in R
Notes:

Response:

***

### Building the Linear Model
Notes:
# in r linear models or lm are in the form: lm(y~x)
# where y = outcome variable &
# x = explanatory variable

# in the I() function below, the I stands for as is
# so in the following example, the I function tells to transform the variable
# for instance the carat variable before using it in the linear model

#before buildig a linear model we can check the fit by using the lmfit function

```{r Building the Linear Model}

#lmfit

m1 <- lm(I(log(diamonds$price)) ~ I(diamonds$carat^(1/3)), data = diamonds)
m1
summary(m1)
anova(m1)
plot(m1)

m2 <- update(m1, ~ . + diamonds$carat)
m3 <- update(m2, ~ . + diamonds$cut)
m4 <- update(m3, ~ . + diamonds$color)
m5 <- update(m4, ~ . + diamonds$clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems
Video Notes:

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

Response:

***

### A Bigger, Better Data Set
Notes:

```{r A Bigger, Better Data Set}

#install.packages('bitops')
#install.packages('RCurl')
library('bitops')
library('RCurl')

#diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
#load(rawConnection(diamondsurl))
diamondsurl = "https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda"
load(url(diamondsurl))
#getwd()
#list.files()
#BigDiamonds <- load("BigDiamonds.Rda")
#colnames(BigDiamonds)
#head(BigDiamonds)
#head(sort(table(BigDiamonds), decreasing =T))
#table(BigDiamonds)
#summary(BigDiamonds)
#describe(BigDiamonds)
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set
Notes:

```{r Building a Model Using the Big Diamonds Data Set}

#m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsbig)
#limiting the dataset by the limiting the features to price < 100000 and diamonds having GIA certification
colnames(diamondsbig)

m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsbig[diamondsbig$price <= 10000 & diamondsbig$cert == "GIA",])
m2 <- update(m1, ~ . + cut)
m3 <- update(m2, ~ . + color)
m4 <- update(m3, ~ . + clarity)
m5 <- update(m4, ~ . + carat)
mtable(m1, m2, m3, m4, m5)
```


***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
#Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
library(memisc)

thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
summary(thisDiamond)

modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)

exp(modelEstimate)
```

Evaluate how well the model predicts the BlueNile diamond's price. Think about the fitted point estimate as well as the 95% CI.

***

## Final Thoughts
Notes:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

