Lesson 5
========================================================

### Multivariate Data
Notes:

***
### adding the correct libraries
getwd()
setwd('C:/Users/akshi/Downloads')
library(rmarkdown)
library(ggplot2)
library(dplyr)
library(ggvis)
library(Hmisc)
library(gridExtra)
library(tidyr)

fb <- read.csv('pseudo_facebook.tsv', sep = '\t')
fb

colnames(fb)


### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
gender <- fb$gender
age <- fb$age
friend_count <- fb$friend_count


```{r Third Qualitative Variable}
### boxplots

ggplot(aes(x = gender, y = age),
       data = subset(fb, !is.na(gender))) + geom_boxplot()

ggplot(aes(x = gender, y = age),
       data = subset(fb, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y = mean, geom = 'point', shape = 4)

ggplot(data = subset(fb, !is.na(gender)),
       aes(x = age, y = friend_count)) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)

### filter function is used to remove the na values in the gender
### summarise function removes just one layer of grouping once
### as we also have the age variable along with the gender i'm adding the 
### ungroup function to remove both the layers
### the arrange function will arrange the data frame by age
fb.fc_by_age_gender <- fb %>%
  filter(!is.na(gender)) %>%
  group_by(age, gender) %>%
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>%
  ungroup() %>%
  arrange(age)

fb.fc_by_age_gender
  
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}

ggplot(data = fb.fc_by_age_gender,
       aes(x = age, y = median_friend_count)) +
  geom_line(aes(color = gender), stat = 'summary', fun.y = median)



```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:
#dcast is used because the desired data format is a dataframe
# we can use acast if the desired data format is an array

```{r}
install.packages('reshape2')
library(reshape2)

fb.fc_by_age_gender.wide <- dcast(fb.fc_by_age_gender,
                                  age ~ gender,
                                  value.var = 'median_friend_count')

fb.fc_by_age_gender.wide
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}

head(fb.fc_by_age_gender.wide)

colnames(fb.fc_by_age_gender.wide)
female.wide <- fb.fc_by_age_gender.wide$female
male.wide <- fb.fc_by_age_gender.wide$male

fb.fc_by_age_gender.wide$ratio <- female.wide/male.wide

colnames(fb.fc_by_age_gender.wide)
head(fb.fc_by_age_gender.wide)

p1 <- ggplot(data = fb.fc_by_age_gender.wide,
             aes(x = female.wide, y = male.wide))

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}

```

***

### Cut a Variable
Notes:

```{r Cut a Variable}

```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}

```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}

```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}

```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

```

***

### Prices over Time
Notes:

```{r Prices over Time}

```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}

```

***

### The Limits of Cross Sectional Data
Notes:

***

### Many Variables
Notes:

***

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

